% !TeX spellcheck = en_US
% !TeX root = ../MS_analysis_thesis.tex

\chapter{Analysis of Selected Biomarkers}\label{chapter: Analysis of Selected Biomarkers}
\lettrine[lhang = 0.4, findent=-60pt, lines=7]{\textbf{
		\initfamily \fontsize{40mm}{40mm} \selectfont I
		\normalfont}}{n}
\Cref{section:Selecting Relevant Biomarkers}, we selected three relevant biomarkers: 2MWT, resting heart rate and sleep duration.
In this chapter, we take a closer look at these selected biomarkers.
For each of these selected biomarkers, we look at the corresponding data and combine this data with the questionnaire data for day ratings to see how useful the biomarker can be in predicting good and bad days.
In particular, we want to relate the selected biomarkers to the `How was your MS today?' question in the `Mijn Kwik' application used by the participants.
This specific question can be answered with `good' (goed), `average' (gemiddeld) or `bad' (slecht).
We will be looking at consecutive days, and how these biomarkers possibly play a role in the quality of upcoming days.
Despite the limited dataset (consisting of data gathered over a period of a couple of weeks), we try to get a better understanding of these biomarkers and how they possibly relate to good and bad days.
As noted before, we will be looking at the data of five participants, each of which is identifiable by a unique identity. 
For the patient that was wearing the Microsoft Band 2, the code to integrate the device's data with our database was not finished while writing this thesis.
Therefore, the data from this participant was not used in any of the analyses of the biomarkers.

\input{sections/2MWT}
\input{sections/RHR}
\input{sections/sleep_duration}
